---
title: "Import AQI data"
author: "Michael Andreae"
date: "Friday, June 19, 2015"
output: pdf_document
---

#Insurance status predicts antiemetic use 
We investigate the Hypothesis that insurance status predicets antiemetic use the population in the Public Use File of the Anestehsia Quality Institute with electronic anesthesia records recording antiemetic use.



```{r, packages, message=FALSE, echo=FALSE, warning=FALSE}
library(knitr) # required to set options in Markdown R
```



```{r, global_options, echo=FALSE}

# set the directory to the project directory
setwd("C:/Users/Micheal/Dropbox/Professional/AQI/AQI_RegressionAnalysis")

# set options
opts_chunk$set(fig.width=6, fig.height=4, fig.path='Figs/',
               echo=FALSE, warning=FALSE, message=FALSE)
```

# Load original dataset and save as Rdata 

```{r, eval=FALSE, echo=TRUE}
# run only once
PUF_Q4_2013_Antimetic <- read.csv("Analysis/Data/PUF_Q4_2013_Antimetic.csv")
save(PUF_Q4_2013_Antimetic, file="Analysis/Data/PUF_Q4_2013_Antimetic.Rdata")
```

# Load Rdata AQI raw dataset

```{r}
rm(list = ls())
load("Analysis/Data/PUF_Q4_2013_Antimetic.Rdata")
```

# Inspect variables

```{r, eval=FALSE, results='hide'}
# name the variable under inspection as myvariable

attach(PUF_Q4_2013_Antimetic)
myvariable <- "patient_age"
myvariable2 <- "procedure_start_time"
unique( get(myvariable) )
class( get(myvariable) )
hist( get(myvariable) )
anyNA(get(myvariable))
table(get(myvariable))
range(get(myvariable))

patient_age[patient_age==-1] <- NA
```

## Predictor insurance status


```{r}
PUF_Q4_2013_Antimetic$Payment[PUF_Q4_2013_Antimetic$Payment == ""] <- NA
PUF_Q4_2013_Antimetic$Payment <- droplevels(PUF_Q4_2013_Antimetic$Payment)
levels(PUF_Q4_2013_Antimetic$Payment)[1] <- "Commercial"
anyNA(PUF_Q4_2013_Antimetic$Payment)
```

The predictor insurance status (*Payment*) is coded in `r length(levels(PUF_Q4_2013_Antimetic$Payment))` levels as `r levels(PUF_Q4_2013_Antimetic$Payment)`, after we removed `r sum(is.na(PUF_Q4_2013_Antimetic$Payment))` cases without insurance information, (orginally coded as ""), these are removed from the dataset.

## Outcome variable antiemetic administration

### Ondansetron

```{r}
PUF_Q4_2013_Antimetic$ondansetron <- 
  as.factor(PUF_Q4_2013_Antimetic$Antiemetics_ondansetron)
levels(PUF_Q4_2013_Antimetic$ondansetron) <- c("not admin", "admin")
```

### Dexamethason

```{r}
PUF_Q4_2013_Antimetic$dexamethason <- 
  as.factor(PUF_Q4_2013_Antimetic$Antiemetics_dexamethason)
levels(PUF_Q4_2013_Antimetic$dexamethason) <- c("not admin", "admin")
```

### Droperidol

```{r}
PUF_Q4_2013_Antimetic$droperidol <- 
  as.factor(PUF_Q4_2013_Antimetic$Antiemetics_droperidol)
levels(PUF_Q4_2013_Antimetic$droperidol) <- c("not admin", "admin")
```

### Ondansetron or Dexamethason

```{r}
ondan_dex_either <- 
  as.factor((PUF_Q4_2013_Antimetic$Antiemetics_ondansetron == 1) | (PUF_Q4_2013_Antimetic$Antiemetics_dexamethason ==1))
levels(ondan_dex_either) <- c("neither", "either")
table(ondan_dex_either)
#table(ondan_dex_either,ondansetron)

head(ondan_dex_either)
head(PUF_Q4_2013_Antimetic$Antiemetics_dexamethason == 1)
```

Both antiemetics ondansetron **OR** dexamethason were sometimes administered together
`r table(ondan_dex_either)` 

This is coded in *ondan_dex_either*

## Potential confounders and other variables

### practice ID versus facility ID

```{r}
attach(PUF_Q4_2013_Antimetic)
kable(table(facilityID,practiceID))
```

The table of facility ID versus practice ID suggests that five practices have only one facility ID and one practice (=5013437) has three (sub) facilities. We will simplify by using practice ID, which has no NA.


### case_duration_minutes

Case duration in minutes (*case_duration_minutes*) is an integer and has `r sum(case_duration_minutes==-1)` missing values coded as -1, which we recoded as NA and removed from the dataset.

```{r, echo=TRUE}
PUF_Q4_2013_Antimetic$case_duration_minutes[PUF_Q4_2013_Antimetic$case_duration_minutes==-1] <- NA
```


### patient age

Patient age (*patient_age*) is an integer with a distribution below and has `r sum(patient_age==-1)` missing values coded as -1, which we recoded as NA and removed from the dataset

```{r, echo=TRUE}
PUF_Q4_2013_Antimetic$patient_age[PUF_Q4_2013_Antimetic$patient_age==-1] <- NA

hist(PUF_Q4_2013_Antimetic$patient_age, 
     main = "Histogram of Age Distribution",
     xlab = "Age in years")
```


### patient_age_group

```{r, echo=TRUE}
levels(PUF_Q4_2013_Antimetic$patient_age_group)[2] <- "1-18"
levels(PUF_Q4_2013_Antimetic$patient_age_group)[1] <- NA
PUF_Q4_2013_Antimetic$patient_age_group <- 
  relevel(PUF_Q4_2013_Antimetic$patient_age_group, "Under 1")
```

Patient age group (*patient_age_group*) is a factor with `r length(levels(PUF_Q4_2013_Antimetic$patient_age_group))` levels: `r levels(PUF_Q4_2013_Antimetic$patient_age_group)`; it has `r sum(is.na(PUF_Q4_2013_Antimetic$patient_age_group))` missing values initially coded as -1, which we recoded as NA and removed from the dataset. (We corrected the miscoding from "18-Jan"" to "1-18").

### patient_sex

```{r}
PUF_Q4_2013_Antimetic$patient_sex <- 
  as.factor(PUF_Q4_2013_Antimetic$patient_sex)

levels(PUF_Q4_2013_Antimetic$patient_sex)[1] <- NA
levels(PUF_Q4_2013_Antimetic$patient_sex)[1] <- "female"
levels(PUF_Q4_2013_Antimetic$patient_sex)[2] <- "male"
```

Patient gender (*patient_sex*) is recoded as factor with the two levels `r levels(PUF_Q4_2013_Antimetic$patient_sex)` and `r sum(is.na(PUF_Q4_2013_Antimetic$patient_sex))` NAs, which are removed from the dataset.

### in_or_out_patient

```{r}
PUF_Q4_2013_Antimetic$in_or_out_patient <- 
  as.factor(PUF_Q4_2013_Antimetic$in_or_out_patient)

levels(PUF_Q4_2013_Antimetic$in_or_out_patient)[1] <- NA
levels(PUF_Q4_2013_Antimetic$in_or_out_patient)[1] <- "Outpatient"
levels(PUF_Q4_2013_Antimetic$in_or_out_patient)[2] <- "Inpatient"
```


in- or outpatient status (*in_or_out_patient*) is recoded as a factor with the two levels `r levels(PUF_Q4_2013_Antimetic$in_or_out_patient)` and `r sum(is.na(PUF_Q4_2013_Antimetic$in_or_out_patient))` NAs, which are too numerous to include this variable in the dataset.

# Clean AQI_4_14 dataframe saved as Rdata

```{r}

attach(PUF_Q4_2013_Antimetic)
myAQI_4_14 <- data.frame(Payment, 
                         ondansetron,
                         dexamethason,
                         droperidol,
                         ondan_dex_either,
                         practiceID=as.factor(practiceID), 
                         case_duration_minutes, 
                         patient_age, patient_age_group, 
                         patient_sex)
myAQI_4_14 <- myAQI_4_14[complete.cases(myAQI_4_14),]
save (myAQI_4_14, file="Analysis/Data/myAQI_4_14.Rdata")
str(myAQI_4_14)