---
title: "Logistic regression AQI data"
author: "Michael Andreae"
date: "July, 14, 2015"
output: pdf_document
---

# Insurance status predicts antiemetic use 
We investigate the Hypothesis that insurance status predicets antiemetic use the population in the Public Use File of the Anestehsia Quality Institute with electronic anesthesia records recording antiemetic use.



```{r, packages, message=FALSE, echo=FALSE, warning=FALSE}
library(knitr) # required to set options in Markdown R
# library(lme4)
# library(nlme)
```



```{r, global_options, echo=FALSE}

# set the directory to the project directory
setwd("C:/Users/Micheal/Dropbox/Professional/AQI/AQI_RegressionAnalysis")

# set options
opts_chunk$set(fig.width=6, fig.height=4, fig.path='Figs/',
               echo=FALSE, warning=FALSE, message=FALSE)
```

## Load cleaned dataset *fullAQI_4_14* and *prov1_AQI_4_14*
 
we load the cleaned larger dataset without and with provider information *fullAQI_4_14.Rdata*, *prov1_AQI_4_14*, which we generated in *import_AQI_14Jul2015.Rmd*   


```{r, echo=TRUE}
rm(list = ls())
load("Analysis/Data/fullAQI_4_14.Rdata")
load("Analysis/Data/prov1_AQI_4_14.Rdata")
str(fullAQI_4_14)
str(prov1_AQI_4_14)

```

## Logistic Regression

```{r}
library(lme4)
```

### Logistic Model 1  Link: logit

We fit a logistic regression model with the a **logit** link.


```{r, echo=TRUE}
formula <- ond ~ pay +age_group +sex +
           ASA +anes_type +duration +practice  

fit_logistic <- glm(formula, 
                   family = binomial(link = "logit"), 
                   data = fullAQI_4_14)
```


```{r, echo=FALSE}
p_values <- summary.glm((fit_logistic))$coefficient[,4]

OR <- exp(coef(fit_logistic)) # exponentiated coefficients
CI_fitlogistic<- confint(fit_logistic) # 95% CI for the coefficients
kable(data.frame(OR, exp(CI_fitlogistic), 
                 p_values), digits = 3)

#summary(fit_logistic) # display results
#exp(confint(fit_logistic)) # 95% CI for exponentiated coefficients
#predict(fit_logistic, type="response") # predicted values
#residuals(fit_logistic, type="deviance") # residuals
```

Summary:
Controlling for age, sex, facilty and case duration, antiemetic administration is strongly associated with insurance status as a marker of SES, but the OR is not as strong as previously estimated controlling for fewer variables. it makes sense that controlling for likely confounders like anesthesia type and ASA status reduces the effect estimate. 

