---
title: "Logistic regression AQI data"
author: "Michael Andreae"
date: "July, 14, 2015"
output: pdf_document
---

# Insurance status predicts antiemetic use 
We investigate the Hypothesis that insurance status predicets antiemetic use the population in the Public Use File of the Anestehsia Quality Institute with electronic anesthesia records recording antiemetic use.



```{r, packages, message=FALSE, echo=FALSE, warning=FALSE}
library(knitr) # required to set options in Markdown R
# library(lme4)
# library(nlme)
```



```{r, global_options, echo=FALSE}

# set the directory to the project directory
setwd("C:/Users/Micheal/Dropbox/Professional/AQI/AQI_RegressionAnalysis")

# set options
opts_chunk$set(fig.width=6, fig.height=4, fig.path='Figs/',
               echo=FALSE, warning=FALSE, message=FALSE)
```

## Load cleaned dataset *fullAQI_4_14* and *prov1_AQI_4_14*
 
we load the cleaned larger dataset without and with provider information *fullAQI_4_14.Rdata*, *prov1_AQI_4_14*, which we generated in *import_AQI_14Jul2015.Rmd*   


```{r, echo=TRUE}
rm(list = ls())
load("Analysis/Data/fullAQI_4_14.Rdata")
load("Analysis/Data/prov1_AQI_4_14.Rdata")
str(fullAQI_4_14)
str(prov1_AQI_4_14)

```

## Logistic Regression

```{r}
library(lme4)
```

### Logistic Model 1  Link: logit

We first fit a logistic regression model with the a **logit** link, saved in *LogisticRegression_AQIFixedEfffects_LogitLink* as shown below:


```{r, echo=TRUE,eval=FALSE}
formula <- ond ~ pay +age_group +sex +ASA +anes_type +duration +practice  

fit_logit <- glm(formula, 
                   family = binomial(link = "logit"), 
                   data = fullAQI_4_14)
```


```{r, echo=FALSE, eval=FALSE}
p_values <- summary.glm((fit_logistic))$coefficient[,4]

OR <- exp(coef(fit_logistic)) # exponentiated coefficients
CI_fitlogistic<- confint(fit_logistic) # 95% CI for the coefficients

kable(data.frame(OR, p_values), digits = 3)

#summary(fit_logit) # display results
#exp(confint(fit_logistic)) # 95% CI for exponentiated coefficients
#predict(fit_logistic, type="response") # predicted values
#residuals(fit_logistic, type="deviance") # residuals
```

Summary:
Controlling for age, sex, facilty and case duration, antiemetic administration is strongly associated with insurance status as a marker of SES, but the OR is not as strong as previously estimated controlling for fewer variables. it makes sense that controlling for likely confounders like anesthesia type and ASA status reduces the effect estimate. 


### Logistic Model 2 Link : log

We try to fit a logistic regression model with the a **log** link.

##### Problem: 
We are unable to fit a log link model with all predictors, because we get the error message:

*Error: no valid set of coefficients has been found: please supply starting values*

if we use more predictors than a few predictors. Two solutions:

* supply starting values for all coefficients, (factors need one for each level)
* consider using the package *biglm*, specifically for log link with large data set

### Starting values for log link glm

Estimate fit with **one less** predictor:

```{r, echo=TRUE}

formula0 <- ond ~ pay +age_group +sex

fit_log0 <- glm(formula0, 
             family = binomial(link = "log"), 
             data = fullAQI_4_14)
```

Use estimated coefficients as **starting values** for model with one additional predictor. Starting values for additional predictors are set to zero.

```{r}
formula <- ond ~ pay +age_group +sex +ASA # one additonal predictor ASA

fit_log <- glm(formula, 
             family = binomial(link = "log"), 
             data = fullAQI_4_14, start = c(coef(fit_log0),rep(0,5) ))
# start = c( coef(fit_log0), rep(0,5) ) ### For each level of ASA one start value!
```

```{r}
p_values <- summary.glm((fit_log))$coefficient[,4]
OR <- exp(coef(fit_log)) # exponentiated coefficients
CI_fitlog<- confint(fit_log) # 95% CI for the coefficients
kable(data.frame(OR, exp(CI_fitlog), 
                 p_values), digits = 3)
```

