---
title: "LogLink Logistic regression AQI data"
author: "Michael Andreae"
date: "July, 24, 2015"
output: pdf_document
---

# Insurance status predicts antiemetic use 
We investigate the Hypothesis that insurance status predicets antiemetic use the population in the Public Use File of the Anestehsia Quality Institute with electronic anesthesia records recording antiemetic use.



```{r, packages, message=FALSE, echo=FALSE, warning=FALSE}
library(knitr) # required to set options in Markdown R
# library(lme4)
# library(nlme)
```



```{r, global_options, echo=FALSE}

# set the directory to the project directory
setwd("C:/Users/Micheal/Dropbox/Professional/AQI/AQI_RegressionAnalysis")

# set options
opts_chunk$set(fig.width=6, fig.height=4, fig.path='Figs/',
               echo=FALSE, warning=FALSE, message=FALSE)
```

## Load cleaned dataset *fullAQI_4_14* and *prov1_AQI_4_14*
 
we load the cleaned larger dataset without and with provider information *fullAQI_4_14.Rdata*, *prov1_AQI_4_14*, which we generated in *import_AQI_14Jul2015.Rmd*   


```{r, load_clean_dataset, echo=TRUE}
rm(list = ls())
load("Analysis/Data/fullAQI_4_14.Rdata")
load("Analysis/Data/prov1_AQI_4_14.Rdata")
str(fullAQI_4_14)
str(prov1_AQI_4_14)

```

## Logistic Model 2 Link : log with glm

We try to fit a logistic regression model with the a **log** link with the 

formula <- ond ~ pay +age_group +sex +ASA +anes_type +duration +practice  

### Problem: 
We are unable to fit a log link model with all predictors, because we get the error message:

*Error: no valid set of coefficients has been found: please supply starting values*

if we use more predictors than a few predictors. Two solutions:

* supply starting values for all coefficients, (factors need one for each level)
* consider using the package *biglm*, specifically for log link with large data set

## Building Loglink with starting values 

Use estimated coefficients as **starting values** for increasingly complex model with one additional predictor per step. Starting values for additional predictors are set to zero.

### Estimate fit_log0 with **four** predictors: 

formula0 <- ond ~ pay +age_group +sex

```{r, log0, echo=TRUE}

formula0 <- ond ~ pay +age_group +sex

fit_log0 <- glm(formula0, 
             family = binomial(link = "log"), 
             data = fullAQI_4_14)
```


```{r log0_results, eval=TRUE}
p_values <- summary.glm((fit_log0))$coefficient[,4]
OR <- exp(coef(fit_log0)) # exponentiated coefficients 
# CI_fitlog<- confint(fit_log) # 95% CI for the coefficients, takes too long
kable(data.frame(OR, p_values), digits = 3, caption = "Results log0")
```

### Estimate fit_log1 with **five** predictors: 

formula1 <- ond ~ pay +age_group +sex +ASA

```{r, log1, echo=TRUE}
formula1 <- ond ~ pay +age_group +sex +ASA # one additonal predictor ASA

fit_log1 <- glm(formula1, 
             family = binomial(link = "log"), 
             data = fullAQI_4_14, start = c(coef(fit_log0),rep(0,5) ))
# start = c( coef(fit_log0), rep(0,5) ) ### For each level of ASA one start value!
```

```{r log1_results, eval=TRUE}
p_values <- summary.glm((fit_log1))$coefficient[,4]
OR <- exp(coef(fit_log1)) # exponentiated coefficients 
# CI_fitlog<- confint(fit_log) # 95% CI for the coefficients, takes too long
kable(data.frame(OR, p_values), digits = 3 , caption = "Results log1")
```

### Estimate fit_log2 with **six** predictors: 

formula2 <- ond ~ pay +age_group +sex +ASA +anes_type 

```{r, log2, echo=TRUE}
formula2 <- ond ~ pay +age_group +sex +ASA +anes_type 

fit_log2 <- glm(formula2, 
             family = binomial(link = "log"), 
             data = fullAQI_4_14, start = c(coef(fit_log1),rep(0,5) ))
```

```{r log2_results, eval=TRUE}
p_values <- summary.glm((fit_log2))$coefficient[,4]
OR <- exp(coef(fit_log2)) # exponentiated coefficients 
# CI_fitlog<- confint(fit_log) # 95% CI for the coefficients, takes too long
kable(data.frame(OR, p_values), digits = 3, , caption = "Results log2")
```

### Estimate fit_log3 with **seven** predictors: 

formula3 <- ond ~ pay +age_group +sex +ASA +anes_type +duration

```{r, log3, echo=TRUE}
formula3 <- ond ~ pay +age_group +sex +ASA +anes_type +duration

fit_log3 <- glm(formula3, 
             family = binomial(link = "log"), 
             data = fullAQI_4_14, start = c(coef(fit_log2),rep(0,1) ))
```

```{r log3_results, eval=TRUE}
p_values <- summary.glm((fit_log3))$coefficient[,4]
OR <- exp(coef(fit_log3)) # exponentiated coefficients 
# CI_fitlog<- confint(fit_log) # 95% CI for the coefficients, takes too long
kable(data.frame(OR, p_values), digits = 3, caption = "Results log3")
```

### Estimate fit_log4 with **seven** predictors: 

formula4 <- ond ~ pay +age_group +sex +ASA +anes_type +duration +practice

```{r, log4, echo=TRUE, eval=FALSE}
formula4 <- ond ~ pay +age_group +sex +ASA +anes_type +duration +practice

fit_log4 <- glm(formula4, 
             family = binomial(link = "log"), 
             data = fullAQI_4_14, start = c(coef(fit_log3), rep (0,3) ) )
```

**Non convergence**
*Warning messages:
1: step size truncated due to divergence 
2: step size truncated due to divergence 
3: step size truncated due to divergence 
4: step size truncated due to divergence 
5: glm.fit: algorithm did not converge*

```{r eval=FALSE}
p_values <- summary.glm((fit_log))$coefficient[,4]
OR <- exp(coef(fit_log)) # exponentiated coefficients
CI_fitlog<- confint(fit_log) # 95% CI for the coefficients
kable(data.frame(OR, exp(CI_fitlog), 
                 p_values), digits = 3)
```

